---
title: "Beals_Brandi Programming with R Test #1"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
# DO NOT ADD OR REVISE CODE HERE
knitr::opts_chunk$set(echo = TRUE)
```

### Test Items

##### (1) (3 points) This problem deals with vector manipulations.

(1)(a)  Create a vector that contains the following, in this order, and print the contents.  Do not round off any values unless requested.
* A sequence of integers from 5 to 10, inclusive.
* A two repetitions of the vector c(2, -5.1, -23).
* The value of the sum of 7/42 and 3

```{r test1a, eval = TRUE, echo = TRUE}
vector <- c(seq(5,10),rep(c(2,-5.1,-23),2),(7/42)+3)
vector
```

(1)(b) Extract the first and last elements of the vector you have created to form another vector using the extracted elements. Form a third vector from the elements not extracted.  Print these vectors.

```{r test1b, eval = TRUE, echo = TRUE}
combo <- c(vector[1],vector[length(vector)])
combo
other <- vector[3:length(vector)-1]
other
```

(1)(c) Use the vectors from (b) to reconstruct the original vector in (a).  Print this vector. Sum the elements and round to two decimal places. Show this result.

```{r test1c, eval = TRUE, echo = TRUE}
original <- c(combo[1],other,combo[2])
original
round(sum(original),2)
```

-----

##### (2) (6 points) The expression y = sin(x) + cos(x) is a trigonometric function.

(2)(a) Using the trigonometric function above, write a function in R that accepts values for x and returns a value for y.

```{r test2a, eval = TRUE, echo = TRUE}
brandi <- function(x){
  y <- sin(x) + cos(x)
  y
}
```

(2)(b) Create a vector, x, of 4001 equally-spaced values from -2 to 2, inclusive. Compute values for y using the vector x and your function in (a).  **Do not print x or y.**  Find the value in the vector x that corresponds to the maximum value in the vector y. In other words, restrict attention to only the values of x and y you have computed. Round to 3 decimal places and print the value of x you find and the corresponding maximum value for y.

Finding the two desired values can be accomplished in as few as two lines of code.  Do not use packages or programs you may find on the internet or elsewhere. Do not print the elements of the vectors x and y. Use coding methods shown in the *Quick Start Guide for R*.

```{r test2b, eval = TRUE, echo = TRUE}
x <- seq(from = -2.0, to = 2.0, length.out = 4001)
y <- brandi(x)
round(x[which.max(y)],3)
round(y[which.max(y)],3)
```

(2)(c) Plot y versus x in color, with x on the horizontal axis. Show the location of the maximum value of y determined in 2(c).  Add a title and other features such as a legend.

```{r test2c, eval = TRUE, echo = TRUE}
plot(x,y,type="l",col="blue",main="Range of Values and Resulting Function Results",sub="Maximum Point is Highlighted")
points(x[which.max(y)],y[which.max(y)],pch=15,col="red")
legend("bottomright","( x , sin(x)+cos(x) )")
```

-----

##### (3) (8 points) Use the "trees" dataset for the following items. This dataset has three variables (Girth, Height, Volume) on 31 trees.

(3)(a) Use *data(trees)* to load the file.  Check the structure with *str()*. Use *apply()* to return the median values for the three variables in "trees." Using R and logicals, give the row number and print the three measurements - Girth, Height and Volume - of the tree with Volume equal to the median of Volume.

```{r test3a, eval = TRUE, echo = TRUE}
data(trees)
str(trees)
apply(trees,2,median)
location <- trees$Volume == median(trees$Volume)
trees[location,]
```

(3)(b) Girth is defined as the diameter of a tree taken at 4 feet 6 inches from the ground. Convert each diameter to a radius, r. Calculate the cross-sectional area of each tree using pi times the squared radius. Present stem-and-leaf plots of the radius and area. 

```{r test3b, eval = TRUE, echo = TRUE}
r <- trees$Girth/2
a <- pi*r^2
stem(r)
stem(a)
```

(3)(c) Use par(mfrow = c(1, 4)) and present colored boxplots of the radii and areas calculated in (b) along with Height and Volume. Label each accordingly.

```{r test3c, eval = TRUE, echo = TRUE}
par(mfrow=c(1,4))
boxplot(r,col="green")
title("Radius")
boxplot(a,col="orange")
title("Girth Area")
boxplot(trees$Height,col="purple")
title("Height")
boxplot(trees$Volume,col="yellow")
title("Volume")
```

(3)(d) Demonstrate that the outlier revealed in the boxplot of Volume is not an extreme outlier.  It is possible to do this with one line of code using *boxplot.stats* or logicals.

```{r test3d, eval = TRUE, echo = TRUE}
boxplot.stats(trees$Volume,coef=3)$out
```

-----

##### (4) (2 points) Use matrix operations shown in the "Quick Start Guide" to solve the following system of linear equations. Display the R script and the numerical solutions for x, y and z. Use matrix operations with your solution to reproduce the values 1, 1, 3 as a means of checking if your solution is correct.  This last demonstration can be accomplished with matrix multiplication in one line of code.  

**x - y + z = 1**

**x - y - z = 1**

**x + y - z = 3**

```{r test4, eval = TRUE, echo = TRUE}
matrix <- matrix(c(1,-1,1,1,-1,-1,1,1,-1),ncol=3,byrow=T,dimnames=list(c("1","2","3"),c("x","y","z")))
matrix
results <- c(1,1,3)
results
solution <- solve(matrix) %*% results
solution
matrix %*% solution
```

-----

##### (5) (6 points) The Cauchy distribution is an example of a "heavy-tailed" distribution in that it will have outliers in both tails.  This problem involves comparing it with a normal distribution which typically has very few outliers.  

5(a) Use *set.seed(124)* and *rcauchy()* with n = 100, location = 0 and scale = 0.1 to generate a random sample designated as y.  Generate a second random sample designated as x with *set.seed(127)* and *rnorm()* using n = 100, mean = 0 and sd = 0.15.  

Generate a new object using *cbind(x, y)*. Do not print this object.  Use *apply()* with this object to compute the inter-quartile range or IQR for each column:  x and y.  Round the results to two decimal places and present.  (The point of this exercise is to demonstrate the similarity of the IQR values.)

For information about rcauchy(), use help in R (?rcauchy).  **Do not print x and y.**

```{r test5a, eval = TRUE, echo = TRUE}
set.seed(124)
y <- rcauchy(100,location=0,scale=0.1)
set.seed(127)
x <- rnorm(100,mean=0,sd=0.15)
combo <- cbind(x,y)
round(apply(combo,2,IQR),2)
```

(5)(b) This item will illustrate the difference between a heavy tailed distribution and one which does not have heavy tails.  Use *par(mfrow = c(2, 2))* and present two histograms in color, one for x and the second for y. Set *xlim = c(-7, 7)* for the Cauchy results, and *xlim = c(-0.45, 0.45)* for the Normal results.  Use the same intervals respectively for ylim with the boxplots.  

```{r test5b, eval = TRUE, echo = TRUE}
par(mfrow=c(2,2))
hist(x,xlim=c(-0.45,0.45),ylim=c(0,50),col="grey")
hist(y,xlim=c(-7,7),ylim=c(0,50),col="pink")
```

(5)(c) QQ plots are useful for detecting the presence of heavy tailed distributions. Use *par(1,2)* and present side-by-side plots, one for each sample, using *qqnorn() and qqline()*.  Add color and titles.   Use cex = 0.5 to control the size of the plotted data points.

```{r test5c, eval = TRUE, echo = TRUE}
par(1,2,mfrow=c(1,2))
qqnorm(x,col="grey",main="Quantile-Quantile Plot of X",cex=0.5,pch=16)
qqline(x)
qqnorm(y,col="pink",main="Quantile-Quantile Plot of Y",cex=0.5,pch=16)
qqline(y)
```